Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Intro nits

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization stack safety that we are aware of is
recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  But our formulations of these properties are
rather different from those of Skorstengaard et al.  Instead of
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we define it
\emph{explicitly} using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity},
\emph{stepwise stack confidentiality}, and \emph{return to caller}.
%
A key technical novelty in the definitions of confidentiality and integrity,
compared to standard formulations from the security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of {\em return to caller}, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:control}).
This reflects a cleaner division between data protection and control-flow protection
than offered by LSE and WBCF.

@@

\label{sec:controlflow}

>>>>

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of stack safety that we are aware of is
recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  But our formulations of these properties are
rather different from those of Skorstengaard et al.  Instead of
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we define it
\emph{explicitly} using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity},
\emph{stepwise stack confidentiality}, and \emph{return to caller}.
%
A key technical novelty in the definitions of confidentiality and integrity,
compared to standard formulations from the security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of {\em return to caller}, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:control}).
\rb{\cref{sec:RTC}?}
This reflects a cleaner division between data protection and control-flow protection
than offered by LSE and WBCF.

@@

\label{sec:controlflow} % RB: ???
